### 2016-11-02 ms.tex
Fix punctuation in App

%\documentclass[preprint,12pt]{aastex}
\documentclass[twocolumn]{emulateapj}
%\usepackage{apjfonts}
\usepackage{color}
\usepackage{amsmath}

\newcommand{\apjvec}[1]{\mbox{\boldmath{$#1$}}}
\newcommand{\apjmat}[1]{{\mathbf{#1}}}

\renewcommand{\theenumi}{\roman{enumi}.}
\renewcommand{\labelenumi}{\theenumi}

\newcommand{\vx}{\apjvec{x}}
\newcommand{\vf}{\apjvec{f}}
\newcommand{\vxi}{\apjvec{\xi}}
\newcommand{\vt}{\apjvec{\theta}}
%\newcommand{\mf}{\apjmat{F}}
\newcommand{\snIa}{\mbox{SN\,{\sc{}i}a}}

\newcommand{\erfc}{{\rm{}erfc}}

\newcommand{\rr}{\mathbf{r}}
\newcommand{\hh}{\mathbf{h}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\ww}{\mathbf{w}}
\newcommand\avg[1]{\langle{#1}\rangle}

\newcommand*{\defeq}{\stackrel{\text{def}}{=}}

% % % % % % % % % % %
% Macros added by TL:

% Misc symbols:
\newcommand{\scond}[1]{\qquad||\;{#1}}  % Skilling conditional
\newcommand{\ctxt}{\mathcal{C}}  % context/background information

% Data and likelihoods:
\newcommand{\eind}{\epsilon}  % epoch index
\newcommand{\edata}{D_\eind}  % epoch-specific image data
\newcommand{\like}{\mathcal{L}}  % generic likelihood
\newcommand{\slf}{\mathcal{L}}  % source likelihood
\newcommand{\flike}{\ell}  % flux likelihood factor
\newcommand{\dlike}{m}  % drxn likelihood factor
\newcommand{\mlike}{\mathcal{M}}  % marginal likelihood

% Source properties & characteristics:
%\newcommand{\schar}{\chi}  % source characteristics (to "characterize" a source)
%\newcommand{\sobs}{\mathcal{O}}  % source observables
\newcommand{\flux}{f}
\newcommand{\fest}{\hat{\flux}}  % flux estimate
\newcommand{\fsig}{\sigma}  % flux uncertainty
\newcommand{\drxn}{\mathbf{n}}
\newcommand{\dest}{\hat{\mathbf{n}}}  % drxn estimate
\newcommand{\dsig}{\delta}  % drxn uncertainty

% Detection, nondetection:
\newcommand{\dtxn}{\mathcal{D}}  % detection
\newcommand{\ndtxn}{\mathcal{N}}  % non-detection
\newcommand{\accept}{\mathcal{A}}  % set of detection data
\newcommand{\fth}{\flux_{\rm th}}

\newcommand{\npd}{\nu}  % noise peak dist'n

% Macro for an "editorial note"; uncomment the 2nd definition to
% produce a version without notes (put it in main doc if desired).
\newcommand\enote[1]{ %
\marginpar[\raggedleft\large $\blacktriangleright$]%
{\raggedright\large $\blacktriangleleft$} %
{\large $\langle\langle\langle$}{\sl #1}{\large  $\rangle\rangle\rangle$} %
}
%\renewcommand\enote[1]{\relax}

% Marginal note (just a redef of \marginpar or \marginparodd).
% Uncomment the 2nd definition to produce a version without notes (put it 
% in main doc if desired).
% Could use \footnotesize for a bigger font.
% The \hspace enables hyphenation of the 1st word.
\setlength{\marginparwidth}{.7in}
\newcommand\mnote[1]{\-\marginpar[\raggedleft\tiny\hspace*{0pt}#1]%
{\raggedright\tiny\hspace*{0pt}#1}}
%\renewcommand\mnote[1]{\relax}

% % % % % % % % % % %

%================================================================================
\begin{document}

\title{Incremental Faint Object Detection in Multi-Epoch Observations}
%\journalinfo{To appear in ApJ?}
\journalinfo{Submitted to Apj}

\author{ Tam\'{a}s Budav\'{a}ri, Alexander S.\ Szalay}
\affil{Dept.\ of Physics and Astronomy, The Johns Hopkins University, 3400 North Charles Street, Baltimore, MD 21218}
\and
\author{Thomas J. Loredo}
\affil{Center for Radiophysics and Space Research, Cornell University, Ithaca, NY 14853}

\shortauthors{}
\shorttitle{}

\begin{abstract} 
Observational astronomy in the time-domain era faces several new challenges. 
One of them is the efficient use of observations obtained at multiple epochs. 
The work presented here addresses faint object detection, and seeks to find an incremental strategy for separating real objects from artifacts in ongoing surveys, in situations where the available data are summaries of the full image data, such as source flux and direction estimates (with uncertainties) reported in catalogs.
We adopt a Bayesian approach, addressing object detection by calculating the probabilities for hypotheses asserting there is no object, or one object, in a small image patch containing at most one detected source at each epoch.
The object-present hypothesis interprets the sources in a patch as arising from a genuine object; the no-object hypothesis interprets any detected sources as spurious, arising from noise peaks.
We study the detection probability for constant-flux objects in a simplified Gaussian noise setting, comparing results based on single exposures and stacked exposures to results based on catalog summary data.
Computing the detection probability based on catalog data amounts to generalized cross-matching: it is the product of a factor accounting for matching of the estimated fluxes of candidate sources, and a factor accounting for matching of their estimated directions (i.e., directional cross-matching across catalogs).
We find that probabilistic fusion of multi-epoch catalog information can detect sources with only modest sacrifice in sensitivity compared to stacking.
We show that the probabilistic cross-matching underlying our approach plays an important role in maintaining detection sensitivity.
\end{abstract}

\keywords{methods: statistical --- surveys --- catalogs --- photometry --- astrometry}

%================================================================================
\section{Introduction}
\label{sec:intro}
\noindent
%
In the era of time-domain survey astronomy, dedicated telescopes scan the sky every night and strategically revisit the same area several times.
The raw data are images, but surveys commonly provide, not only image data, but also \emph{catalogs}, summaries of the image data that aim to enable a wide variety of studies without requiring users to analyze raw or processed image data.
Catalogs typically report object properties, based on algorithms that detect sources in images with a measure of statistical significance above some threshold, chosen so that the resulting catalog is likely to be highly pure (i.e., with few or no spurious sources).

The question we address in this paper is how to combine information from a sequence of independent observations to maximize the ability to detect faint objects at or near a chosen detection threshold, ameliorating the data explosion due to false detections from a lower threshold that would be required by a suboptimal method.
Focusing on the faint objects that typically dominate the collected data is an important and timely problem for a number of ongoing surveys and vital for planning the next-generation data processing pipelines.

There are two different ways one can approach the problem. 
A traditional, resource-intensive approach is to wait until all observations are completed, performing detection by stacking the multi-epoch image data.
An optimal procedure for threshold-based detection with image stacks was introduced by \citet{chisq}.
Once a master object catalog is produced from the stacked images, time-series of source measurements are created by forced photometry at the master catalog locations.

An alternative (non-exclusive) approach is to perform source detection independently for each observation, producing a catalog of candidate sources at each epoch.
The detection threshold may be different for each epoch. 
Interim object catalogs may be produced by analysis of the series of overlapping source detections potentially associated with a single object using any available data; a final catalog would be built using the catalogs from all epochs.
Of course, a catalog based on data from many epochs should be able to include many dim sources that escape confident detection in single-epoch or few-epoch catalogs.
To enable construction of a deep multi-epoch catalog, the single-epoch catalogs must report information for candidate sources with relatively low statistical significance; i.e., the single-epoch catalogs must have reduced purity.
If we set the single-epoch threshold too high, there will be too few detections; we will not have well-sampled time series for dim sources, and the final catalog will be too small.
If, on the other hand, we set the threshold too low, the single-epoch catalogs will be overwhelmed with (mostly false) detections that were seen only once, requiring wasteful expenditure of storage and computing resources for constructing multi-epoch catalogs.
An optimal threshold might preserve the size and quality of the final catalog, while enabling users to build interim catalogs on-the-fly, potentially tailored to specific, evolving needs.

We here address the second alternative, considering how best to use the list of detections while the observations are in progress to prune spurious source detections but keep the sources associated with genuine objects.
The study presented here is exploratory, to establish the basic ideas and provide initial metrics for studying the feasibility of the incremental approach.
To make the analysis analytically tractable and the results easy to interpret, we restrict ourselves to an idealized setting; we will present a more general treatment in a subsequent paper.

%================================================================================
\section{Detection Probabilities} 
\label{sec:det}

\noindent

We adopt the terminology of LSST and other time-domain synoptic surveys, using \emph{source} to refer to single-epoch detection and measurement results, and \emph{object} to refer to a unique underlying physical system (e.g., a star or galaxy) that may be associated with one or more sources.
(Here we limit ourselves to objects that would appear as a single source.)
For simplicity, we consider observations in a single band unless stated otherwise.

Consider an object with constant flux $f$ and direction $\drxn$ (a unit-norm vector on the sky). 
At each epoch $\eind$, analysis of the image data $\edata$ corresponding to a small patch of sky of solid angle $\Omega$ produces a \emph{source likelihood function} (SLF) for the basic observables, flux $\flux$ and direction $\drxn$, of a candidate source in the patch.
The SLF is the probability for the data as a function of the (uncertain) values of the observables,
\begin{equation}
\like_\eind(\flux,\drxn) \equiv p(\edata|\flux,\drxn,\ctxt),
\label{like-def}
\end{equation}
where $\ctxt$ denotes various contextual assumptions influencing the analysis,
e.g., specification of the photometric model and information about instrumental and sky backgrounds.
(Since $\ctxt$ is common to all subsequent probabilities, we henceforth consider it as implicitly present.)
For example, if photometry is done via point spread function (PSF) fitting with weighted least squares, and if the noise level and backgrounds are known, then it may be a good approximation to take $\like_\eind(\flux,\drxn) \propto \exp[-\chi^2(\flux, \drxn)/2]$, where $\chi^2(\flux, \drxn)$ is the familiar sum of squared weighted residuals as a function of the source flux and direction.

We consider a catalog at a given epoch to report summaries of the likelihood functions for candidate sources that have met some detection criteria.
The most commonly reported summaries are best-fit fluxes (or magnitudes) with a quantification of flux uncertainty (typically a standard error or the half-width of a 68\% confidence region), and, separately, best-fit sky coordinates with an uncertainty for each coordinate.%
%***\footnote{Cautionary note about the meaning of ``best-fit'' here.}
We here take these summaries to correspond to a factored approximation of the source likelihood function,
\begin{align} \label{eq:eplike}
\like_\eind(\flux, \drxn)
  &\equiv p(\edata|\flux, \drxn, H_1)\nonumber\\
  &= \flike_\eind(\flux)\, \dlike_\eind(\drxn),
\end{align}
where the epoch-specific flux factor, $\flike_\eind(\flux)$, is a Gaussian with mode $\fest_\eind$ (the catalog flux estimate at epoch $\eind$) and standard deviation $\fsig_\eind$, and the direction factor, $\dlike_\eind(\drxn)$, is an azimuthally symmetric bivariate Gaussian with mode $\dest_\eind$, and standard deviation $\dsig_\eind$.
This may be a rather crude approximation; we will address it further elsewhere, here merely noting that it is implicitly adopted for most survey catalogs.
For simplicity, we take the flux factors to have the same standard deviation at all epochs,  $\fsig_\eind = \sigma$.
\enote{Define the direction uncertainty, too.}

% These symbols unused?

%Let $\dtxn_\eind$ denote a proposition asserting that a candidate source was detected in the patch at epoch $\eind$, i.e., that $\edata$ passes detection criteria.\footnote{We are assuming that detection is deterministic, i.e., given the full data $\edata$, detection is a yes-no proposition.}
%$\dtxn_\eind$ may be expressed more explicitly as $\edata \in \accept_\eind$, asserting that the data from epoch $\eind$ lie in the acceptance set, $\accept_\eind$, the possible values of $D_\eind$ classified as corresponding to a detected source.
%Let  $\ndtxn_\eind$ denote the proposition stating that no candidate source was detected in the patch at epoch $\eind$.
%Equivalently, we may write $D_\eind\notin \accept_\eind$.
%We here adopt the simple criterion that $\hat{\flux}_\eind$ be greater than a fixed threshold, $\fth$, common across all epochs.

We adopt a simple source detection criterion: a candidate source with a flux likelihood mode $\fest$ larger than a single-epoch threshold value, $\fth$, is deemed a detection.
The probability for detection in a single-epoch catalog is the probability that source with true flux $f$ will produce a single-epoch measurement $\fest_\eind$ that falls above the threshold.
This probability is just the integral of the Gaussian flux likelihood function above the threshold, which we denote by
%
\begin{equation}\label{eq:pf}
P_f \defeq P(>\!\fth|f) = \frac{1}{2}\,\erfc\left(\frac{\fth\!-\!f}{\sigma\sqrt{2}}\right),
\end{equation}
%
where $\erfc(\cdot)$ is the complementary error function.

For comparison, consider detection probabilities in the case of stacked exposures from $k$ observations.
We assume that the objects are stationary and have a constant flux, and that the dominant source of noise is still the sky, so the relative noise is reduced by $\sqrt{k}$ after stacking.
For a stacked exposure flux threshold $f_S$, the probability for detection is
%
\begin{equation}
P'_f \defeq \frac{1}{2}\,\erfc\left(\frac{f_S\!-\!f}{\sigma\sqrt{2/k}}\right).
\end{equation}
%
Figure~\ref{fig:1} displays the detection probability as a function of true object flux for single-epoch and stacked data, for various choices of the single-epoch and stacked thresholds.
The dotted green lines represent the single-exposure situation $P_f$ as a function of the true flux in $\sigma$ units for detection thresholds of 2, 3, 4, and 5$\sigma$.
Similarly the solid yellow lines correspond to the stacked detections with \mbox{$k\!=\!9$} exposures.

Consider two curves corresponding to the same threshold, so $f_S = \fth$.
The probability for detection at $f=\fth$ is 50\% for both a single exposure and stacked exposures.
But the curve for stacked exposures is much steeper, with a higher probability for detecting sources brighter than $\fth$, and a lower probability for detecting sources dimmer than $\fth$.
That is, when constructed with a common threshold, the catalog built from stacked data will be more complete above threshold, and will more effectively exclude sources with true flux below threshold.

\begin{figure}
\epsscale{1.2}
\plotone{fig/f1.png}
\caption{The detection probability is shown in different scenarios as a function of the true flux in $\sigma$ units. The green dotted lines illustrate single-exposure cases with different thresholds that take values of 2, 3, 4, and 5$\sigma$ ({\it{}from left to right}). The yellow solid lines are the same for stacks with \mbox{$k\!=\!9$} observations, at the same flux thresholds.}
\label{fig:1}
\end{figure}

%--------------------------------------------------------------------------------
\subsection{Multiple detections and non-detections}
\noindent
%
Faint sources will not always be detected.
%
The probability for making $n$ detections among \mbox{$k=n+m$} observations follows the binomial distribution, giving the multi-epoch detection probability,
%
\begin{equation} \label{eq:binomial}
 P(n|k,f) = {k \choose n}\ P_f^n\,\big(1\!-\!P_f\big)^{k-n}.
\end{equation}
%
An interesting quantity is the probability that an object would lead to source detections in $n_0$ or more observations.
This is simply the sum
\begin{equation}
P(n\!\geq{}\!n_0|k,f) = \sum_{n=n_0}^k {k \choose n}\ P_f^n\,\big(1\!-\!P_f\big)^{k-n}
\end{equation}
%
(this can be expressed in terms of the incomplete beta function).
In Figure~\ref{fig:2} we plot these probabilities as a function of $f$ (again in $\sigma$ units) for \mbox{$k$=9} observations.
From left to right, the solid red curves show the probability for detecting an object of given flux in exactly 1, 2, etc., up to 9 observations.
Similarly the dashed blue curves correspond to cases $1+$ (1 or more), $2+$, and so on.
(Note that the functions coincide for the case $n=n_0=k=9$.)

\begin{figure}
\epsscale{1.2}
\plotone{fig/f2.png}
\caption{The detection probability in multiple observations is shown here as function of the true flux. Assuming \mbox{$k\!=\!9$} total available observations, the {\it{}solid red curves} show the probability of the object appearing in exactly 1, 2, 3, etc., up to 9 observations ({\it{}from left to right}). Similarly, the {\it{}dashed blue lines} correspond to the 1+, i.e., 1 or more, 2+, 3+, etc., cases.}
\label{fig:2}
\end{figure}

Figure~\ref{fig:3} compares detection probability curves for the stacked exposure case (solid yellow curves, as in Figure~\ref{fig:1}) and the multi-epoch, $n>n_0$ detection case (dashed blue curves, as in Figure~\ref{fig:2}).
For a particular stacked exposure case, we see there is a multi-epoch case whose detection probability curve displays very similar performance.
For example, the $3\sigma$ stacked exposure curve is very similar to the multi-epoch $5+$ detection case.
This indicates that collecting sources with $5+$ detections from \emph{single-epoch} $3\sigma$ catalogs is nearly equivalent in terms of catalog completeness and purity to producing a separate, new $3\sigma$ stacked exposure catalog.
%
\enote{Move the rest to intro?}
Analyzing the single-epoch catalogs has a number of advantages.
It can be implemented in an incremental fashion that follows the schedule of the survey and the time-series is readily available at a given time; there is no need to go back to old images and to performed forced photometry at locations that are revealed in the stacks only.

\begin{figure}[t]
\epsscale{1.2}
\plotone{fig/f3.png}
\caption{The comparison of the probabilities plotted previously in Figures~\ref{fig:1} and \ref{fig:2} reveals the similarities of the alternative methods. The curves of the stacked cases ({\it{}solid yellow lines}) and the summed up binomials ({\it{}dashed blue lines}) follow similar trends in the usual regime of parameter space. In particular, we highlight the remarkable agreement of the 3$\sigma$ curve for the stack detections ({\it{}solid yellow line second from the left}) and the 5+ sum of the binomials ({\it{}dashed blue in the middle}).}
\label{fig:3}
\end{figure}

%================================================================================
\section{Distinguishing Real Sources from Noise Peaks}
\noindent
%
The previous calculations addressed detectability of a source of known true flux, $f$.
In real-life scenarios, the problem is quite the opposite---we are presented with the observations and would like to understand the properties of the sources. 
In this context, our focus is on how one can reliably distinguish noise peaks from real sources. 
It is important to emphasize that we have more information than just the fact that a source has been detected; we also have flux measurements, at multiple epochs. 
Our approach is motivated by Bayesian hypothesis testing, where the strength of evidence for presence of a source is quantified by the posterior probability for the source-present hypothesis, or equivalently, by the posterior odds in favor of a source being present vs.\ being absent (the odds is the ratio of probabilities for the rival hypotheses).
The posterior odds is the product of prior odds and the data-dependent \emph{Bayes factor}.
The prior odds depends on population properties; it may be specified a priori when there is sufficient knowledge of the population under study, or learned adaptively by using hierarchical Bayesian methods (for examples of this in the related context of cross-identification, see REF).
\enote{Get SCMA5 ref.}
Here we focus on the Bayes factor; we will address hierarchical modeling in a follow-up paper.

\enote{Should ``source'' be ``object'' here?}

The Bayes factor is the ratio of marginal likelihoods for the competing hypotheses, one that claims that the source is real, and its complement that assumes just noise:
%
\begin{equation} \label{eq:Bfac}
B = \frac{\mlike_{\rm{}real}}{\mlike_{\rm{}noise}} .
\end{equation}
%
Each marginal likelihood, $\mlike_{\rm{}hyp}$, is the integral, with respect to all free parameters for the hypothesis, of the product of the likelihood function and the prior probability density for the parameters.

Let us now assume that out of $k$ observations, we measure $n$ detections with measured fluxes $\{\fest_\eind\}$.
We consider the two competing hypotheses separately.

%--------------------------------------------------------------------------------
\subsection{Real-source hypothesis}
\noindent
Let $\dtxn$ denote the set of indices for epochs with detections, and $\ndtxn$ denote the set of indices for epochs with nondetections:
\begin{equation}
\begin{split}
\dtxn  &\equiv \{\eind : \fest_\eind > \fth\},\\
\ndtxn &\equiv \{\eind : \fest_\eind \le \fth\}.
\end{split}
\label{D-ND-sets}
\end{equation}
For a candidate source with $n$ detections among $k$ catalogs, the likelihood for a candidate true flux $f$ is
%
\begin{equation}
\like(f) = (1\!-\!P_f)^{k-n}\,\prod_{\eind \in \dtxn} \flike_\eind(\flux),
\end{equation}
%
where \mbox{$(1\!-\!P_f)$} is the probability of not detecting an object with true flux $f$, which happens \mbox{$(k\!-\!n)$} times, and $\flike_\eind(\cdot)$ is the flux likelihood function defined above (Gaussians with means equal to $\fest_\eind$).
%
The marginal likelihood for the real-source hypothesis is obtained by averaging over all possible true flux values, $f$.
For a source that is a member of a population with known flux probability density $\pi(f)$, the prior probability for $f$, used for the averaging in the marginal likelihood, is $\pi(f)$, so that
%
\begin{equation}
\mlike_{\rm{}real} = \int\!\!df\,\pi(f)\,L(f)
\end{equation}
%
which is a one-dimensional integral that can be analytically or numerically quickly evaluated.
(When the population distribution is not known a priori, it may be estimated via joint analysis of the catalog data for many candidate sources, within a hierarchical model, a significant complication that we will elaborate on elsewhere.)

%--------------------------------------------------------------------------------
\subsection{Noise peak hypothesis}\label{sec:peaks}
\noindent
%
The alternative hypothesis is that the detections are simply random noise peaks in the image.
The noise hypothesis marginal likelihood, $\mlike_{\rm noise}$, is the probability for the catalog data presuming no real source is present.\footnote{We are presuming the use of a fixed source detection algorithm, e.g., fixed apertures. 
If the algorithm is adaptive, its tuning parameters need to be accounted for in this probability.}

For epochs with a candidate source reported in the catalog, the datum is the flux measurement, $\fest_\eind$, and the relevant factor in the marginal likelihood is $\npd(\fest_\eind) \defeq p(\fest_\eind|\rm{Noise})$, the \emph{noise peak distribution}, evaluated at $\fest_\eind$.
This distribution will depend on the noise statistics for each catalog.

For epochs with no reported detection, we instead know only that $\fest_\eind \le \fth$, so the relevant factor is the fraction of \emph{missed} noise peaks,%
\footnote{The integral is over possible \emph{measured} flux values, $\fest_\eind$, not true flux; in the Gaussian regime assumed here the measured value may be negative, albeit with small probability.
The estimated flux would be constrained to be positive via the prior density, $\pi(f)$, which would multiply the flux likelihood when computing posterior flux estimates.}
\begin{equation} \label{eq:Q-noise}
Q_N \defeq \int_{-\infty}^{\fth}\!\!d\fest\,\npd(\fest).
\end{equation}
The probability for a false detection is then $P_N = 1 - Q_N$.

To compute these quantities comprising $\mlike_{\rm noise}$, we need to know the noise peak distribution, $\npd(\fest)$.
This distribution is not trivial to specify; it will depend both on the noise sources, and on the source detection algorithm.
%
Typically, a source finder performs a scan, identifying local peaks of the measured fluxes smoothed with a kernel, e.g., corresponding to a specified point source aperture.
Under the noise hypothesis, the source finder will be finding peaks of a smooth random field.
The locations and amplitudes of the peaks will form a point process, whose statistical properties can be analytically calculated \citep{adler,bbks,bond,kaiser}.
%
The most important consequence is that even though the underlying noise at the pixel level may be independent and Gaussian, the source finder output will correspond to sampling from a point process with a more complicated distribution of fluxes.
In particular, although the pixel-level noise distributions are symmetric (about the mean background), the distribution for (falsely) detected fluxes is skewed toward positive values.
%
The relevant calculation is presented in the Appendix.

Figure~\ref{fig:surface} shows the surface density of noise peaks as a function of the detection statistic $\fest$ (in $\sigma$ units), in the scenario when the sky noise is spatially independent and Gaussian.
The surface density is in units of objects per $a^2$, where $a$ is the width of the point spread function (see Appendix).
The noise peak distribution is the normalized version of this function.
The surface density has a mode at $\fest \approx 1.33\sigma$ and its shape is well approximated with a Gaussian with standard deviation $\approx 0.835$ for all positive values of $\fest$. 
The shaded (magenta) area highlights the excess density over the Gaussian at negative $\fest$ values.

\begin{figure}
\epsscale{1.2}
\plotone{fig/f4.png}
\caption{The thick (magenta) curve illustrates the surface density of noise peaks as a function of flux in the scenario when the sky noise is white, i.e., has a flat spectrum. The mode of the distribution is at around 1.33$\sigma$ (vertical line) and its shape is well approximated with a Gaussian (thin gray line) for all positive fluxes. The excess density over the Gaussian at negative flux values is shaded (magenta) for clarity.}
\label{fig:surface}
\end{figure}

As noted above, we obtain the probability for detecting a noise peak, $P_N$, by integrating $\npd(\fest)$ above the flux threshold.
%
Figure~\ref{fig:frac}a shows the results as a function of the flux threshold in $\sigma$ units (left), and on a scale corresponding to an LSST-like magnitude (right).
\enote{We should specify how the LSST magnitude scale was set, or point to a REF.}
We see that the fraction of noise peaks above threshold is about 62\% at 1$\sigma$, dropping quickly to about a few percent at 3$\sigma$, and becoming negligible at 5$\sigma$.
%
Based on just this figure, it is tempting to set a high detection threshold to reject such ``ghost peaks'' and keep the catalog of detections nearly pure; but that would mean we lose the opportunity to recover the numerous really faint sources.

\begin{figure*}
\epsscale{1.2}
\plotone{fig/f5.png}
\caption{The fraction of detected noise peaks drops quickly by raising the threshold. At 1$\sigma$ the value is about 62\% but at 3$\sigma$ it is only a few percent and at 5$\sigma$, which is 24 magnitudes in this case, the detected fraction is negligible. Naively this makes it highly desirable to put a harder constraint on the detection limit but then the opportunity is lost to track fainter sources. The right solution is not this shortcut.}
\label{fig:frac}
\end{figure*}

%\begin{figure*}
%\epsscale{1.2}
%\plotone{fig/f6.png}
%\caption{Panels (a)-(d) show the distribution of noise peaks and galaxies as a function of flux and magnitude using linear and logarithmic scales. The solid (magenta) line shows the distribution of the noise peaks, which is the normalized version of the curve in Figure~\ref{fig:surface}. For comparison, the dotted (green) Gaussian represents the distribution of measurements in randomly placed apertures. The dashed (black) line illustrates the \mbox{$1/f^2$} galaxy distribution used in this study.}
%\label{fig:distr}
%\end{figure*}

Our multi-epoch approach suggests a different strategy: instead of seeking to make the catalogs for \emph{each} epoch pure, we can adopt a lower single-epoch threshold, relying on the fusion of data across epochs to weed out ghosts.
The marginal likelihood and Bayes factor computations accomplish this data fusion.

%For a specific detection threshold we can calculate the noise peak detection probability $P_N$ as described above; e.g., 1$\sigma$ yields \mbox{$P_{N}\!\simeq\!0.62$}. 
The marginal likelihood for the noise hypothesis is a product of the terms for the detections and non-detections:
%
\begin{equation}
\mlike_{\rm noise} = \left(1\!-\!P_N\right)^{k-n}\,\prod_{\eind \in \dtxn} \npd(\fest_\eind).
\end{equation}
%
We now have all the ingredients for computing the Bayes factor of Eq.~\ref{eq:Bfac}, providing an objective measure of how much the data prefer a real-source origin to a noise peak origin.

%--------------------------------------------------------------------------------
\subsection{Displacements: Including the Astrometry}
\label{sec:astrom}
\noindent
So far we have only used the flux information in the data. 
Genuine sources should have both consistent fluxes and consistent directions across all epochs.
In practice, due to the noise and astrometric errors, the detections of the same object will shift in each exposure, thus the resulting catalogs have to be cross-matched. 
Using a probabilistic method can be to our direct benefit here, enabling straightforward combination of the flux and direction information.
%%
%Bayesian cross-identification \citep{pxid} compares the marginal likelihood of the hypothesis that all detections of a set belong to the same source against the marginal likelihood under the opposite assumption.
%%
%Every candidate association is assigned a Bayes factor $B_{\rm{}pos}$ based on the celestial positions of its detections.

The detections from a real source are all connected, they are just displaced by a random astrometric error; but noise peaks (ghosts) will be independent of each other and their associations can only be by chance.
As we are working under the approximation that the flux and sky position estimates are independent (see Eq.~\ref{eq:eplike}), the Bayes factor using both the photometric and astrometric information factors,
%
\begin{equation}
B_{\rm{}flux,pos} = B_{\rm{}flux}\cdot{}B_{\rm{}pos}.
\end{equation}
%
\enote{Find REF.}
The astrometric cross-match Bayes factor, $B_{\rm{}pos}$, has been derived in \ref{XX} (see Eq.~(17) there, and Eq.~(19) for the tangent plane Gaussian limit that holds for high-precision astrometry).
That work also discusses generalizations that account for proper motion and other complications.

In the following section we assess the discriminative power of multi-epoch source detection by applying it to simulated galaxies and noise peaks, both omitting and including the astrometric data.

\enote{It might be interesting to show astrometry-only results. They might correspond to allowing for arbitrary variability, so that no flux matching is expected at all. Use of variability models would then produce results between the pure astrometry and astrometry $+$ flux cases.}

%================================================================================
%\section{Caveats}
%\label{sec:caveats}

%================================================================================
\section{Simulations}
\label{sec:disc}

\noindent
We here use simulations to demonstrate the detection capability of our multi-epoch approach in a setting with known ground truth.
The simulation parameters were chosen to produce data similar to that provided by modern large-scale optical surveys.

%--------------------------------------------------------------------------------
\subsection{Simulated Galaxies}
\noindent
We assume that galaxies are brighter than 28 magnitudes and that the 5$\sigma$ detection limit is 24 magnitudes, corresponding roughly to parameters of LSST photometry.
%
Panel~(b) of Figure~\ref{fig:frac} shows the noise peak detection probability as a function of magnitude based on these parameters, in contrast to the dimensionless presentation in panel~(a).
%
To compute the marginal likelihood for the source-present hypothesis, we must specify a prior for the source flux, $\flux$.
Here we use a standard faint-galaxy number counts law,
with the number counts following the empirical formula of 
\mbox{$dN\!\propto\!10^{0.4m}dm$}; see \cite{MT00-NumCounts}. 
That approximately translates to the properly normalized population distribution of
%
\begin{equation} 
\label{eq:ref3}
\pi(f) = \left\{\begin{array}{l l}
           f_L/f^2  & \quad \mbox{if\ $f > f_L$}\\
           0 & \quad \mbox{otherwise}
           \\ \end{array} \right.
\end{equation}
%
where $f_L$ is the limiting flux that corresponds to the previously defined magnitude limit.

%\subsection{The Mock Catalogs}\label{sec:mock} \noindent
We generate sets of random detections for 20,000 galaxies with true fluxes between 28 and 23 magnitudes by simply drawing $\fest_\eind$ values from a Gaussian centered on the actual fluxes.
%
We also generate 2,000 ghost detection $\fest_\eind$ values from $\npd(\fest)$ by inverting its cumulative distribution (computed numerically on a grid).
%
The number of exposures is set to the previously used $k=9$ with a flux threshold of just 1$\sigma$, deep in the noise.
In observations with our specified parameters, the number of ghost detections will greatly outnumber the galaxy detections with this low threshold.
The numbers of galaxies and ghosts were chosen to enable display of the distributions of Bayes factors for the two classes of detections (noise and true).

We first analyze the simulated data considering only the photometric information (i.e., ignoring the directional Bayes factors).
In Figure~\ref{fig:bf} the (red) points represent the resulting Bayes factors for the real sources (right of the double dashed vertical lines) and the noise peaks (on the left).
%
The weight of evidence is strong for the bright galaxies but weakens for the faint galaxies.
The smallest Bayes factors arise for galaxies with true magnitudes near 26.5, which corresponds to the mode of the noise peak distribution, $\npd(\fest)$.
Perhaps surprisingly, sources dimmer than magnitude 26.5 can have larger Bayes factors than those with magnitude 26.5.
This happens because $\npd(\fest)$ peaks away from $\fest=0$, i.e., we do not expect noise peaks to have arbitrarily small measured fluxes.
For the weakest detectable sources, the most likely number of detections among the $k=9$ epochs is one.
The flat top of the distribution at the faint end corresponds to very dim sources detected only once, very near threshold.
The smaller Bayes factors in that region of the plot correspond to unlikely larger numbers of detections near the threshold; the discreteness in the number of detections produces a subtle banding in the distribution.

We now additionally consider the astrometric data.
Assuming a constant direction  uncertainty of $0.1\arcsec$ for all detections, we simulate the coordinates for the mock galaxies.
Around the true direction of each object, we randomly generate points from a 2D Gaussian.
This flat sky approximation is excellent in this regime; for such tight scatters, the approximation error is below the limit of the numerical representation of double precision floating point numbers.

The coordinates of noise peaks are generated homogeneously.
%
The surface density of the ghosts is analytically calculated and its integral above the 1$\sigma$ detection threshold yields \mbox{$\nu\!=\!0.04/\square\arcsec$}.
%
A simple algorithm is to pick a large enough square, with area $\Omega$, and randomly draw the number of peaks from a Poisson distribution with expectation value $\lambda=\nu\Omega$.
Out of these ghosts, we pick a number equal to the number of flux detections, with locations such that are closest to the center, where the simulated object is placed.

Figure~\ref{fig:bf2} shows the distribution of Bayes factors, for the real (mock) and noise sources, now including the astrometric cross-matching factor.
Banding due to discreteness in the number of detections is now clearly apparent among the true-object Bayes factors; the 9 levels correspond to the different number of detections with the lowest being 1.
For the lowest band, corresponding to a single detection, $B_{\rm{}pos}$ is unity by definition (no constraint coming from a single detection), producing the same Bayes factor distribution as in the flux-only calculation.
For multiple detections, the Bayes factors for the true sources are greatly enhanced by including astrometric information (note that the ordinate is logarithmic); just two detections produces quite strong evidence for the true-object hypothesis.
The Bayes factors for the noise peaks have moved to much lower values, due to the low likelihood of directional coincidences.

This computation demonstrates that flux and astrometric catalog data, combined across epochs, can strongly distinguish real objects from spurious detections.
We have not addressed what threshold Bayes factor to use for producing a multi-epoch catalog, or, in Bayesian terms, how to convert Bayes factors into posterior probabilities for candidate objects.
As noted above, when the object population density (on the sky and in flux) is known a priori, the calculation is straightforward (e.g., the prior odds will be proportional to the ratio of true object and noise peak sky densities).
When these quantities are unknown, a possibly complicated hierarchical Bayesian calculation can jointly estimate them and the object properties.
When many objects are detected, an approximate approach, plugging in empirical estimates of the densities based on the data, is likely to suffice, as described in \cite{REF,REF}.
\enote{SCMA5 refs?}

%****************************************
\color{red}
\begin{figure}
\epsscale{1.2}
\plotone{fig/f7u-U.png}
\caption{The (red) points represent the weight of evidence for simulated galaxies as a function of their true brightness and the same for generated random noise peaks. The Bayes factor is high for the bright galaxies but its logarithm is close to zero for the faint ones. The noise peaks (left of double lines) seem to separate slightly from the faintest sources (just right of the double dashed vertical lines) because the fluxes in noise peaks tend to be higher, which better mimics sources between 26 and 25 magnitudes.}
\label{fig:bf}
\end{figure}

\begin{figure}[t!]
\epsscale{1.2}
\plotone{fig/f8u-U.png}
\caption{The Bayes factor can differentiate significantly better between real objects and noise peaks when we include astrometric information. Here we plot logarithm of product of the photometric and astrometric Bayes factors, the latter which comes directly from the cross-identification based on the celestial coordinates, see text for details.}
\label{fig:bf2}
\end{figure}

% Removed "mixed detections" and "variable objects" sections here...

%================================================================================
\section{Summary}
\label{sec:sum}
\noindent
In this paper we explored new approaches to processing a series of exposures with special focus on the faintest sources, where most of the new discoveries are expected.
%
We found that one can recover sources just by collecting a number of detections above even a low threshold with the same probability as if the sources were extracted from coadded images. Consequently we can develop incremental algorithms that crossmatch the detections and weed out the noise on the fly from a master source catalog.

We derived the spatial properties of noise peaks that commonly appear in catalogs.
The surface density of these ghosts is asymmetric and skewed toward positive flux values but can be accurately approximated by a shifted Gaussian for most practical purposes.
We used our analytic results to perform Bayesian hypothesis testing to distinguish between true sources and noise peaks.
%
Based on the Bayes factor, the bright sources over \mbox{3$\sigma$} \mbox{($m\!\simeq\!24.55$)} start to separate out from the noise peaks, which becomes with very convincing evidence at \mbox{5$\sigma$} (24 mags).
%
Just based on the flux measurements, the faint sources are practically indistinguishable from the noise peaks but their celestial coordinates can help to correctly identify them as real sources. Using probabilistic cross-identification is to our direct benefit directly providing the required quantities.
For best results we combine the brightness and spatial information in a probabilistic manner by multiplying the Bayes factors.

In general, the specificity and the selectivity of the proposed discriminator depends on a number of parameters most of which we discussed as part of the simulated case study.
The pixel size, the detection limit, the number of exposures and the point-spread function will determine the frequency of noise peaks as well as the statistics of the real sources based on which one can come up with reliable thresholds for progressively weeding out noise from the catalogs throughout the lifetime of a survey.

\acknowledgements{}
The authors gratefully acknowledge valuable and inspiring discussions with Andy Connolly and Robert Lupton on various aspects of the topic.
This study was supported by the ... {\color{red}JHU funding agencies???} and by the NSF via grant AST-1312903.

\color{black}

%================================================================================
\begin{thebibliography}{}

\bibitem[Adler(1981)]{adler} Adler, R.~J.\ 1981, The Geometry of Random Fields, Chichester: Wiley, 1981,

\bibitem[Bardeen et al.(1986)]{bbks} Bardeen, J.~M., Bond, J.~R., Kaiser, N., \& Szalay, A.~S.\ 1986, \apj, 304, 15

\bibitem[Bond \& Efstathiou(1987)]{bond} Bond, J.~R., \& Efstathiou, G.\ 1987, \mnras, 226, 655

\bibitem[Budav{\'a}ri \& Szalay(2008)]{pxid} Budav{\'a}ri, T., \& Szalay, A.~S.\ 2008, \apj, 679, 301

\bibitem[Budav{\'a}ri(2011)]{2011ApJ...736..155B} Budav{\'a}ri, T.\ 2011, 
\apj, 736, 155

\bibitem[Gregory \& Loredo(1992)]{gregory} Gregory, P.~C., \& Loredo, T.~J.\ 1992, \apj, 398, 146

%\bibitem[Kaiser(1984)]{1984ApJ...284L...9K} Kaiser, N.\ 1984, \apjl, 284, L9

\bibitem[Kaiser(2004)]{kaiser} Kaiser, N.\ 2004, ``The Likelihood of Point Sources in Pixellated Images'', Pan-STARRS internal report, PSDC-002-010-xx

% SNANA: A Public Software Package for Supernova Analysis
\bibitem[Kessler et al.(2009)]{snana} Kessler, R., Bernstein, J.~P., Cinabro, D., et al.\ 2009, \pasp, 121, 1028

\bibitem[Loredo(2012)]{loredo} Loredo, T.~J.\ 2012, arXiv:1206.4278

\bibitem[Lund \& Rudemo(2000)]{lund} Lund, J., \& Rudemo, M.\ 2000, Biometrika, 87, 2, pp.235-249 (http://www.jstor.org/stable/2673461)

\bibitem[Kerekes et al.(2010)]{pmxid} Kerekes, G., Budav{\'a}ri, T., Csabai, I., Connolly, A.~J., \& Szalay, A.~S.\ 2010, \apj, 719, 59

\bibitem[Madau \& Thompson(2000)]{MT00-NumCounts} Madau, P., \& Thompson, C.\ 2000, \apj, 534, 239

\bibitem[Press(1997)]{press} Press, W.~H.\ 1997, Unsolved Problems in Astrophysics, p.49-60, arXiv:astro-ph/9604126

% Using Type IA supernova light curve shapes to measure the Hubble constant
\bibitem[Riess et al.(1995)]{riess95} Riess, A.~G., Press, W.~H., \& Kirshner, R.~P.\ 1995, \apjl, 438, L17

\bibitem[Szalay et al.(1999)]{chisq} Szalay, A.~S., Connolly, A.~J., \& Szokoly, G.~P.\ 1999, \aj, 117, 68

\end{thebibliography}

\appendix
\color{blue}

\section{Peaks of Two-Dimensional Random Fields}
\noindent
Consider a two dimensional Gauusian random field $f(\rr)$, with a known power 
spectrum. Its gradient would be $\hh$, and the second derivative tensor $g$. 
We would like to find out the density of peaks of this field above a certain 
height. We will follow the procedure outlined in \citet{bbks}.

We will expand the field and its gradient to second order around a peak at the 
position $\rr_p$:
\begin{eqnarray}
	f(\rr) = f(\rr_p)+\frac{1}{2} g_{ij}(\rr_p) (\rr-\rr_p)_i  (\rr-\rr_p)_j\\
	h_i(\rr) = (\rr-\rr_p)_j g_{ij}(\rr_p),
\end{eqnarray}
where we already use the fact that the gradient of the field at a peak is zero, 
i.e. $h_i(\rr_p)=0$. Provided that $g$ is non-singular at $\rr_p$, we can express 
$\rr-\rr_p$ from the second equation:
\begin{equation}
	\rr-\rr_p = g^{-1}(\rr_p) \hh(\rr_p).
\end{equation}
We can write a Dirac delta that picks all extremal points of $f$ as
\begin{equation}
	\delta^{(2)}(\rr-\rr_p) = |\det g(\rr)| \delta^{(2)}[\hh(\rr)].
\end{equation}
This expression turns a continous random field, defined at all points over our 
two-dimensional space into a discrete point process, that of the extremal points 
of the field,
\begin{equation}
	n_{\rm{}ext}(\rr) = |\det g(\rr)| \delta^{(2)}[\hh(\rr)].
\end{equation}

In order to pick the peaks of the Gaussian random field we will also need to have 
a negative definite $g$. If we only want peaks of a certain height, we need to 
calculate the appropriate ensemble average of this density over the constrained 
range of the variables.

We have six random variables, the field $f$, the three components of the symmetric 
$g$ tensor, and the two components of the gradient $\hh$. The correlations can be 
computed in a straight-forward manner, given the power spectrum of the field. The
gradient is uncorrelated with both the field and the second derivatives, due to the
parity of the Fourier representation. Let us denote the correlation matrix of the
field and the Hessian by $C$, and that of the gradient as $H$. Furthermore, let us 
define the different $k$-moments of the power spectrum characterizing the field as
\begin{equation}
	\sigma_n^2 = \frac{1}{(2\pi)^2} \int d^2k\, k^{2n} P(k).
\end{equation}

We can now explicitely write down the correlation matrix $C$ of 
${\bf v} = (f,g_{11},g_{12},g_{22})$ and $H$ for $\hh= (h_1, h2)$, as
\begin{equation}
	 C = \left(
		\begin{array}{cccc}
			\sigma_0^2 & -\sigma_1^2/2 & 0 & -\sigma_1^2/2\\
			-\sigma_1^2/2 & 3\sigma_2^2/8 & 0 & \sigma_2^2/8\\
			  0 & 0 & \sigma_2^2/8  & 0\\
			-\sigma_1^2/2 & \sigma_2^2/8 & 0 & 3\sigma_2^2/8
		 \end{array}\right),
\end{equation}
\begin{equation}
	 H = \left(
		\begin{array}{cccc}
			-\sigma_1^2/2 & 0 \\
			0 & -\sigma_1^2/2
		 \end{array}\right).
\end{equation}
With these we can write the multivariate Gaussian distribution using the inverse 
of the correlation matrix as a product of two independent distributions
\begin{equation}
	dP =\exp\left(-\frac{\vv^T C^{-1}\vv}{2}\right)
  			\frac{d^4 \vv} {(2\pi)^2|C|^{1/2}}
		 \exp\left(-\frac{\hh^T H^{-1} \hh}{2}\right) 
			\frac{d^2 \hh} {2\pi|H|^{1/2}}.
\end{equation}
Before we proceed further, the second derivative tensor can be described more conveniently
with the two eigenvalues $\lambda_1,\lambda_2$ and a rotation angle,
$\phi$, as follows:
\begin{eqnarray}
	g_{11} =& \lambda_1 \cos^2\phi +\lambda_2 \sin^2\phi,\\
	g_{12} =& (\lambda_1-\lambda_2) \sin\phi \cos\phi,\\
	g_{22} =& \lambda_1 \sin^2\phi + \lambda_2 \cos^2\phi.
\end{eqnarray}
For simplicity let us introduce the dimensionless variables 
$x=(\lambda_1+\lambda_2)/\sigma_2$, the trace of the second derivative tensor, 
$y=(\lambda_1-\lambda_2)/\sigma_2$, and $z = f/\sigma_0$. The Jacobian of the 
transformation from $(f,g_{11},g_{12},g_{22})$ to $(z,x,y,\phi)$ is
\begin{equation}
		J = \sigma_0\sigma_2^3 y/2.  
\end{equation} 
Let us also introduce the dimensionless $\gamma$ and the characteristic scale 
$R$ as
\begin{equation}
	\gamma = \frac{\sigma_1^2}{\sigma_0\sigma_2},
		\qquad R^2 = \frac{\sigma_1^2}{\sigma_2^2}.
\end{equation}
The quadratic form containing $\vv$ in the exponent can be written with the new 
variables as
\begin{equation}
	Q = v^T C^{-1} v = \left(
		2 y^2	+\frac{x^2 + 2\gamma x z + z^2}{1-\gamma^2}\right).
\end{equation}
The determinants of $C$ and $H$ are 
\begin{equation}
	|C| = \frac{1}{64}(1-\gamma^2)\sigma_0^2\sigma_2^6,\qquad
	|H| = \frac{1}{4} \sigma_1^2.
\end{equation}
In these variables, the unconstrained probability distribution for 
$(x,y,z,\phi)$ becomes
\begin{equation}
	dP = \frac{4y}{(2\pi)^2 \sqrt{1-\gamma^2}} \exp(-\frac{1}{2}Q) 
		\,dx\ dy\ dz\ d\phi.
\end{equation}
In order to properly handle the symmetries of the problem, we can assume that 
$\lambda_1\geq\lambda_2$. Then still any $(\lambda_1,\lambda_2)$ pair can be 
mapped onto itself by a 180 degree rotation, so the valid range of $\phi$ is 
$(0,\pi)$.  Since none of the terms depend on $\phi$, we can integrate 
over $\phi$, resulting in
\begin{equation}
	dP =  
		 \exp\left[-\frac{x^2 + 2\gamma x z + z^2}{2(1-\gamma^2)}\right] 
			\frac{\,dx\ dz}{2\pi \sqrt{1-\gamma^2}}
		\left( e^{-y^2}\,2y\ dy\right).
\end{equation}
The constraint $\lambda_1>\lambda_2$ maps onto $0<y$. If we perform the 
integration over $y>0$, and $x,z$ over $(-\infty,\infty)$, we get 1, as we 
should, for the unconstrained probability for a general point.

As we introduce the peak constraints, we need to first consider the impact
on the gradient. The constrained probability distribution
is
\begin{equation}
	dw_x = \exp\left(-\frac{\hh^T H^{-1} \hh}{2}\right) 
			|\det g| \delta^{(2)}[\hh]			
			\frac{d^2 \hh} {2\pi|H|^{1/2}}.
\end{equation}
After integrating over $d^2\hh$ we get the extremum weight
\begin{equation}
	w_x = \frac{|\det g|}{2\pi|H|^{1/2}} =\frac{x^2-y^2}{4 \pi R^2}.
\end{equation}
This will multiply the unconstrained probability for the density of extremal 
points of the random field, 
\begin{equation}
		dn_{\rm{}pk}=\left[\frac{(x^2-y^2)y}{2\pi R^2} e^{-y^2}\,dy\right]
		\exp\left[-\frac{x^2 + 2\gamma x z + z^2}{2(1-\gamma^2)}\right] 
			\frac{\,dx\ dz}{2\pi \sqrt{1-\gamma^2}}.
\end{equation}
For a peak both eigenvalues of the second derivative tensor must be negative. 
In the rotated coordinates $(x,y)$, this means that $x<0$, and $0<y<-x$.
We can easily integrate over the allowed range of $y$ next, yielding
\begin{equation}
	\int_0^{-x} dy\, y\/ (x^2-y^2)\, e^{-y^2} 
		=\frac{1}{2}(x^2-1+e^{-x^2}).
\end{equation}
We are left with
\begin{equation}
	dn_{\rm{}pk} = \frac{(-1 + x^2 +e^{-x^2})}{4\pi R^2}  
	\exp\left[-\frac{x^2 + 2 x z \gamma + z^2 }{2 (1-\gamma^2)}\right]
	\frac{\,dx\ dz}{2\pi \sqrt{1-\gamma^2}}.
\end{equation}
Let us introduce the function $B(x)$ as
\begin{equation}
	B(s,b) = \sqrt{\frac{\pi}{b}} \exp\left({\frac{s^2}{2b}}\right)
	\left[1+ {\rm erf}\left( \frac{s}{\sqrt{2b}}\right)\right].
\end{equation}
Evaluating the integral over $-\infty<x\leq 0$ in Mathematica, we obtain
\begin{equation}
 n_{\rm{}pk}(s,\gamma)= 
		%\exp\left(-\frac{s^2}{2\gamma^2}\right) 
		\frac{e^{-\frac{s^2}{2\gamma^2}}}{8\pi^2R^2}
		\Bigg[
			(1-\gamma^2)s +[s^2 - \gamma^2 (1 + s^2)] B(s,1)
			+B(s, 3- 2\gamma^2),
		\Bigg]
\end{equation}
with
\begin{equation}
	s = \frac{\gamma z}{\sqrt{1-\gamma^2}}.
\end{equation}

We need to evaluate the shape parameter $\gamma$. Assume that the window 
function applied to the random field is a Gaussian with a scale $a$,
\begin{equation}
	w(r) = \frac{1}{2\pi a^2} \exp\left(-\frac{r^2}{2 a^2}\right).
\end{equation}
Its Fourier transform is also a Gaussian,
\begin{equation}
	W(k) = \exp\left(-\frac{k^2a^2}{2}\right).
\end{equation}
We model the sky noise as a white noise with a flat spectrum. Thus the
correlations in the measured random field are determined by the window function,
i.e.,
\begin{equation}
	P(k) = A\,|W(k)|^2 = A \exp\left(-k^2a^2\right).
\end{equation}
With this power spectrum it is straightforward to compute the scale and the shape 
parameters as
\begin{equation}
	\gamma^2 = \frac{1}{2}, \qquad R^2 = \frac{a^2}{2}.
\end{equation}

\end{document}